A Systematic Literature Review on Extensions to Role Based Access Control

### 2013-04-17 paper/sections/analysis.tex
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.

RQ2: What are the motivations behind RBAC extensions? \\

@@

RQ3: Do the extension models have corresponding implementations? \\

@@

\begin{table}
\centering
\begin{tabular}{ | c | c | }
\cline{1-2}

@@

\cline{1-2}
\end{tabular}
\caption{Implementation types found and the count of primary sources}
\label{tab:implementations}
\end{table}

@@

RQ4: How are extension to RBAC evaluated theoritically and in practice? \\

@@

RQ5: What domains have RBAC extensions been created for? \\

@@

The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of extension models.  
Further, the authors noted that the categories associated with papers identifying the medical domain was not limited to one or two but cut across
each of the nine categories except for Organization and Resource. 
The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and the requirements therein.
Given the growth of the research and development of medical applications over the past decade this result does not appear to be surprising. However,
the RBAC standard was originally created to reduce cost and increase interoperability - two goals of current regulation around the standardization
of electronic health record systems. The large number of proposed models, and the cross-category result stand in direct opposition of the goals
of both the RBAC standard and current regulations.

@@

RQ6: What commonalities or generalizations exist across all categorizations? \\

>>>>

RQ2: What are the motivations behind extensions to RBAC?

\subsection{Results}

@@

RQ3: Do the extension models have corresponding implementations?

\subsection{Results}

@@

\begin{table}
\centering
\caption{Implementation types found and the count of primary sources}
\begin{tabular}{ | c | c | }
\cline{1-2}

@@

\cline{1-2}
\end{tabular}
\label{tab:implementations}
\end{table}

@@

RQ4: How are extension to RBAC evaluated theoeetically and in practice?

\subsection{Results}

@@

RQ5: What domains have extensions to RBAC been created for?

\subsection{Results}

@@

The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of extension models.  
Further, the authors noted that the categories associated with papers identifying the medical domain was not limited to one or two but cut across
each of the eight categories except for Organization. 
The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and the requirements therein.
Given the growth of the research and development of medical applications over the past decade this result does not appear to be surprising. However,
the RBAC standard was originally created to reduce cost and increase interoperability - two goals of current regulation around the standardization
of electronic health record systems. The large number of proposed models, and the cross-category result stand in direct opposition of the goals
of both the RBAC standard and current regulations.

@@

RQ6: What commonalities or generalizations exist across all categories?

\subsection{Results}

### 2013-04-17 paper/sections/categorization.tex
Committing updates based on comments from reading group.
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.

RQ1: What categories exist within RBAC extensions? \\

During the paper reading phase, the authors noted that categorical themes were present among the extensions. 
For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy directly within the title of the paper and the name of their model. 
Some papers presented this direct prouncement of a theme, while others were less obvious. 
Thus, the authors developed a set of guidelines to aide in determining a set of extension categories based upon observations during the paper reading phase. The guidelines were:

\begin{itemize}
\item Self Declaration - Does the title or name of the model classify itself?
\item Repetition of Phrase - Does the body of the paper present the same phrase repeatedly when discussing their model?
\item Self Assessment - Do the authors of the paper directly identify a quantifier for their paper within the body of the paper?
\item Simple - Can the category be represented by a single quantifier?
\end{itemize}

If we take the previous example above, the paper "Privacy-aware role-based access control" \cite{ni2010privacy} contained "privacy" in
the title and in the name of the model leading to the creation of the Privacy categorization and the subsequent placement of the paper under that category.
By comparison, the paper "An extended RBAC model based on granular logic" \cite{jian2008extended} does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was context based.

The authors offer definitions for each category based on the data extracted from the primary sources and the English definitions for each word.

@@

  \item Constraint (Const): Access control model in this category extends constraints, which are conditional restrictions on permissions of given roles. This constraint is either static and dynamic. For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician. This example describes doctor's permission with conditional restriction, "only when the doctor is assigned as the designated primary care physician".

@@

  \item Resource (Res): Access control model can be extended to handle any system resources (e.g., a file, printer, terminal, database record, etc) in a flexible way.

@@

The 29 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their assumed
category and further, displays the perceived heirarchy of the categories. 
The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and
the Spatio-Temporal category contained more primary sources that the Spatial or Temporal categories alone.

@@

>>>>

RQ1: What categories exist within RBAC extensions?

\subsection{Results}

During the paper reading phase, the authors noted that categorical themes were present among the extensions. 
For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy directly within the title of the paper and the name of their model. 
Some papers presented this direct pronouncement of a theme, while others were less obvious. 
Thus, we developed a set of guidelines to aide in determining a set of extension categories based upon observations during the paper reading phase. 
In developing these guidelines, we attempted to define each category by a single noun-phrase descriptor.
The guidelines were:

\begin{itemize}
\item Model Name - Does the name of the model classify itself?
\item Self Assessment - Do the authors of the paper directly identify a descriptor for their model within the body of the paper?
\item Repetition of Phrase - Does the body of the paper present the same phrase repeatedly when discussing their model?
\end{itemize}

If we take the previous example above, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in
the title and in the name of the model leading to the creation of the Privacy category and the subsequent placement of the paper under that category.
By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was context based.

We offer definitions for each of the 8 observed categories based on the data extracted from the primary sources and the English definitions for each noun-phrase. 
Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Table \ref{tab:categorization}.

@@

  \item Constraint (Const): Access control model in this category extends constraints, which are conditional restrictions on permissions of given roles. This constraint is either static and dynamic. For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician. This example describes doctor's permission with conditional restriction, ``only when the doctor is assigned as the designated primary care physician''.

@@

@@

The 28 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their assumed
category and further, displays the perceived heirarchy of the categories. 
The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and
the Spatio-Temporal category contained more primary sources that the Spatial or Temporal categories alone.

@@

\greybox{We found eight categories that exist within extensions to RBAC: Constraint, Context, 
Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We further found that the other
seven categories are all fall under the category of Context.}

### 2013-04-17 paper/sections/intro.tex
Committing updates based on comments from reading group.
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item RQ1: What categorizations exist within extensions to RBAC?
\item RQ2: What are the motivations behind extensions to RBAC?
\item RQ3: Do the extensions to RBAC have corresponding implementations?
\item RQ4: How are extensions to RBAC evaluated theoritically and in practice?
\item RQ5: What domains have extensions to RBAC been created for?
\item RQ6: What commonalities or generalizations exist across all categorizations?
\end{itemize}

The authors performed a systematic literature review designed to explore the current body of research in the area of extensions to the core RBAC model.  The review we performed yielded 1716, of which 29 were deemed primary sources for inclusion as model extensions to the RBAC standard model.  This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel. For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.  Further, the review provides insight into the state of evaluation, or lack thereof, within the RBAC extension model community.

@@

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item Summarizes current extended RBAC research work and its contributions.
\item Guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
\item Identifies the research challenges in the areas of security policies and suggest a future extension of RBAC.
\item Identifies the deficiencies in RBAC extension evaluation.
\end{itemize}

>>>>

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item RQ1: What categories exist within extensions to RBAC?
\item RQ2: What are the motivations behind extensions to RBAC?
\item RQ3: Do the extensions to RBAC have corresponding implementations?
\item RQ4: How are extensions to RBAC evaluated theoretically and in practice?
\item RQ5: What domains have extensions to RBAC been created for?
\item RQ6: What commonalities or generalizations exist across all categories?
\end{itemize}

The authors performed a systematic literature review designed to explore the current body of research in the area of extensions to the core RBAC model.  The review we performed yielded 1716, of which 28 were deemed primary sources for inclusion as model extensions to the RBAC standard model.  This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel. For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.  Further, the review provides insight into the state of evaluation, or lack thereof, within the RBAC extension model community.

@@

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item Summarizes current research on extensions to RBAC.
\item Guides a direction for a new standard of an extended RBAC model.
\end{itemize}
